---
title: "R Notebook"
output: html_notebook
---

Problem statement: species do not always have constant observer effects - relative biomass can mask effects on non-dominant species.

Goal: test a new method to do multivariate analysis on species catches and compare these results to those derived from alternative multivariate and single species or univariate approaches.  We do this over multiple degrees of species interactions (dominance) and degrees of bias (observer effects). 



```{r packages}
library(tidyverse)
library(FMAtools)  
library(mvobservr)
library(lme4)
library(tweedie)

source("functions/triplet_stats.R")
source("functions/TripletAnalysis.R")
source("functions/getDescriptiveStats.R")
source("functions/runTandFtests.R")
source("functions/runGLMM.R")
source("functions/assign_sig.R")
source("functions/ObserverEffectStats.R")
source("functions/mvglm_obs_dd.R")
source("functions/anova_obs_dd.R")
source("functions/bootAnova_obs_dd.R")
source("functions/qtweedie_obs_dd.R")
```


```{r simulation_setup}
# Set how many populations per scenario to generate
n_samples_per_level <- 100
# Set the Tweedie power parameter (lambda). 1 < p < 2 is typical for biomass.
tweedie_power <- 1.6
# Set the dispersion parameter for the Tweedie distribution.
# Higher values create more variance (further from target BP).
phi <- 3

# Set the desired total biomass for every saved population.
fixed_total_biomass <- 1000000
# Set the number of vessels in the fleet
# Vessel needs to be defined for some of the analyses
nvess <- 20
# Set the number of trips 
ntrips <- 500
# Set target Berger-Parker index (dominance) levels.
target_bp_levels <- seq(0.5, 0.9, by = 0.1)

# total datasets to create
n_samples_per_level * length(target_bp_levels)
```



```{r generatedata}
# for reproducibility
set.seed(123)

trip_sets <- map(rep(1:length(target_bp_levels), n_samples_per_level), ~{
  #set up data frame for trips
  catches <- data.frame(uid = 1:ntrips)
  
  #create catches
  catches$sp_1 <- rtweedie(ntrips, p = tweedie_power, mu = 100*target_bp_levels[.x], phi = phi)
  catches$sp_2 <- rtweedie(ntrips, p = tweedie_power, mu = 100*(1-target_bp_levels[.x]), phi = phi)
  # the 100 acts as a raising factor to get away from a lot of 0s
  
  #standardize catches
  catches <- catches %>%
    mutate(total_biomass = sum(sp_1 + sp_2)) %>%
    mutate(scalar = fixed_total_biomass/total_biomass) %>%
    mutate(across(c(sp_1, sp_2), function(x) x*scalar)) %>%
    select(-scalar, -total_biomass) %>%
    mutate(biomass_total = sp_1 + sp_2) %>% #this is at the trip level, previously was at the fleet level
    mutate(bp_level = target_bp_levels[.x],
           bp_true = sum(sp_1)/sum(biomass_total))
  
  #add vessel randomly
  catches <- catches %>%
    mutate(vessnum = sample(nvess, size=nrow(.), replace=TRUE)) 
  
}, .progress = TRUE)

```


```{r addbias}
# Set target coverage rate
trip_coverage <- 0.25
# Set bias levels for species (change on observed trips; 0 = no bias, -0.25 = 25% reduction)
bias <- c(0, -0.40) #-0.25, -0.10, -0.40

trip_sets_adj <- map(trip_sets, ~{.x %>%
  mutate(obs = rbinom(nrow(.), 1, trip_coverage)) %>%
  mutate(sp_1 = ifelse(obs == 1, sp_1 * (1+bias[1]), sp_1),
         sp_2 = ifelse(obs == 1, sp_2 * (1+bias[2]), sp_2),
         biomass_total = sp_1 + sp_2
         )
}, .progress = TRUE)

map(trip_sets_adj, ~{
  .x %>%
    summarize(sp2 = mean(sp_2), sp2unobs = mean(ifelse(obs==0, sp_2, NA), na.rm=TRUE),
              sp2obs = mean(ifelse(obs==1, sp_2, NA), na.rm=TRUE))
}) %>%
  list_rbind() %>%
  mutate(b = (sp2obs-sp2unobs)/sp2unobs) %>%
  ggplot(aes(x=b)) + geom_histogram()
```


# run tests separately for time reasons
just run on total biomass 

## t and f
```{r}
res_t <- map(trip_sets_adj, ~runTandFtests(.x, "biomass_total"))
res_t <- list_rbind(res_t, names_to = "set")
res_t

res_t %>% 
  mutate(bp_level = rep(target_bp_levels, n_samples_per_level)) %>% 
  group_by(bp_level) %>% 
  summarize(mean(tp < 0.05), mean(Fp < 0.05))

```

## univariate glmm
```{r}
res_g <- map(trip_sets_adj, ~runGLMM(.x, "biomass_total"), .progress = TRUE)
res_g <- list_rbind(res_g, names_to = "set")
res_g

res_g %>% 
  mutate(bp_level = rep(target_bp_levels, n_samples_per_level)) %>% 
  group_by(bp_level) %>% 
  summarize(mean(AICd>2))
```

## triplet
```{r}
res_tt <- map(trip_sets_adj, ~TripletAnalysis(.x, "biomass_total", bootstrap_reps = 1000), .progress = TRUE) #about 5 min for 500
res_tt <- list_rbind(res_tt, names_to = "set")
res_tt

res_tt %>% 
  mutate(bp_level = rep(target_bp_levels, n_samples_per_level)) %>% 
  group_by(bp_level) %>% 
  summarize(mean(KSp < 0.05), mean(ci_lo > 0 | ci_hi < 0))
```

##permanova
```{r}
adon_formula = "Y~factor(obs)"

res_p <- map(trip_sets_adj, ~{
  Y <- .x %>% dplyr::select(starts_with("sp_"))
  adon <- vegan::adonis2(as.formula(adon_formula), data = .x, permutations = 1000, method="euclidean")
  data.frame(metric = "biomass_total", perma_p = adon$`Pr(>F)`[1])
}, .progress=TRUE) #45min for 500
res_p <- list_rbind(res_p, names_to = "set")
res_p

res_p %>% 
  mutate(bp_level = rep(target_bp_levels, n_samples_per_level)) %>% 
  group_by(bp_level) %>% 
  summarize(mean(perma_p<0.05))
```

##mvglm
```{r}
res_m <- imap(1:length(trip_sets_adj), ~{
  print(paste0("***set ", .x, " ", now()))
  trip_sets_adj[[.x]] %>%
      mutate(observed = ifelse(obs==1, 'Y', 'N')) %>%
      pivot_longer(cols=starts_with("sp_"),
                   names_to = 'species', values_to = 'biomass') %>%
      mvglm_obs_dd(block = NULL, add_var = NULL, n_permutations=100) %>%
      pluck("ps") %>%
      rename(biomass_total = p) %>%
      pivot_longer(everything(), names_to = "metric", values_to = "mvglm_p") 
}) #about 5s per permutation
res_m <- list_rbind(res_m, names_to = "set") %>%
  filter(metric == "biomass_total")
res_m

save(res_m, file="output_data/mvglmloop_b40.Rdata")
```


```{r}
load("output_data/mvglmloop1.Rdata")
res_mc <- res_m
load("output_data/mvglmloop2.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
load("output_data/mvglmloop3.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
load("output_data/mvglmloop4.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
load("output_data/mvglmloop5.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
load("output_data/mvglmloop6.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
load("output_data/mvglmloop7.Rdata")
res_m$set <- res_m$set + nrow(res_mc) #increment set due to segmented runs
res_mc <- bind_rows(res_mc, res_m)
res_m <- res_mc

nrow(res_m)
count(res_m, set) %>% filter(n > 1)
```






## combine all tests
```{r}
res_comb <- map(trip_sets_adj, ~{
        .x %>%
          summarize(bp_level = max(bp_level), bp_true = max(bp_true), cov = mean(obs), sp1 = sum(sp_1), sp2 = sum(sp_2))
    }) %>%
  list_rbind(names_to = "set") %>%
  left_join(res_t, by="set") %>%
  left_join(res_g, by="set") %>%
  left_join(res_tt, by="set") %>%
  left_join(res_p, by="set") %>%
  left_join(res_m, by="set")
res_comb

save(res_comb, file="output_data/alltests_2spp_oe40_cov25.Rdata")
```


```{r}
#load("output_data/alltests_2spp_oe25_cov25.Rdata")
#load("output_data/alltests_2spp_oe05_cov25.Rdata")
#load("output_data/alltests_2spp_oe40_cov25.Rdata")
res_comb %>%
  #assign_sig() %>%
  mutate(t_test_sig = ifelse(tp < 0.05, 1, 0),
           f_test_sig = ifelse(Fp < 0.05, 1, 0),
           dAIC_sig = ifelse(glmmconv==1, ifelse(AICd > 2, 1, 0), NA),
           KS_test_sig = ifelse(KSp < 0.05, 1, 0),
           median_test_sig = ifelse(ci_lo > 0 | ci_hi < 0, 1, 0),
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(bp_level) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni. GLMM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  ggplot(aes(x=bp_level, y=pctsig, col=test)) +
  geom_line(linewidth=1) +
  scale_color_brewer(palette = 'Set2') +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="% of simulations with positive (significant) result",
       title = "power of tests on total biomass",
       subtitle = "25% coverage, 25% reduction on catch of species 2 when observed"
  ) +
  scale_x_continuous(labels=scales::label_percent()) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1))
```


```{r}
res_comb %>%
  mutate(sp2_orig = 1000000-sp1, 
         sp2_bias = (sp2-sp2_orig)/sp2_orig,
         .before = "tstat") %>%
  ggplot(aes(x=sp2_bias/0.25)) + geom_histogram()
```


















```{r}
res_comb %>%
  #assign_sig() %>%
  mutate(
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(bp_level) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  ggplot(aes(x=bp_level, y=mvglm_sig/perma_sig)) +
  geom_line() +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="relative power (mvglm/permanova)",
       title = "relative power of tests on total biomass",
       subtitle = "25% coverage, 25% reduction on catch of species 2 when observed"
  ) +
  scale_x_continuous(labels=scales::label_percent())
```





```{r}
bootstrap_size = 100
n_bootstraps = 500

res_comb %>%
  mutate(t_test_sig = ifelse(tp < 0.05, 1, 0),
           f_test_sig = ifelse(Fp < 0.05, 1, 0),
           dAIC_sig = ifelse(glmmconv==1, ifelse(AICd > 2, 1, 0), NA),
           KS_test_sig = ifelse(KSp < 0.05, 1, 0),
           median_test_sig = ifelse(ci_lo > 0 | ci_hi < 0, 1, 0),
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  select(bp_level, ends_with("sig")) %>%
  mutate(bin = as.factor(as.numeric(bp_level))) %>%
  split(~bin) %>%
  map(.x=., .f=~{slice_sample(.x, n=bootstrap_size*n_bootstraps, replace=TRUE) %>%
      mutate(boot = rep(1:n_bootstraps, each=bootstrap_size)) %>%
      group_by(bin, boot) %>%
      summarize(across(ends_with("sig"), mean), 
                .groups="drop")
    }) %>%
  list_rbind() %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni. GLMM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  group_by(bin, test) %>%
  summarize(meansig = mean(pctsig), lo = quantile(pctsig,0.025), hi = quantile(pctsig, 0.975), .groups="drop") %>%
  filter(test %in% c('MV GLM','permanova','uni. GLMM')) %>%
  ggplot(aes(x=as.numeric(bin), y=meansig, ymin=lo, ymax=hi, col=test, fill=test, group=test)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="% of simulations with positive (significant) result",
       title = "power of tests on total biomass",
       subtitle = "25% coverage, 25% reduction on catch of species 2 when observed"
  ) +
  scale_x_continuous(labels=scales::label_percent()) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1)) 
```




```{r}
load("output_data/alltests_2spp_oe25_cov25.Rdata")
res_oe <- res_comb %>% mutate(effsize = 25)
load("output_data/alltests_2spp_oe10_cov25.Rdata")
res_oe <- res_oe %>%
  bind_rows(res_comb %>% mutate(effsize = 10))
load("output_data/alltests_2spp_oe40_cov25.Rdata")
res_oe <- res_oe %>%
  bind_rows(res_comb %>% mutate(effsize = 40))

table(res_oe$effsize, res_oe$bp_level)
```

```{r}
res_oe %>%
  mutate(
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(bp_level, effsize) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  ggplot(aes(x=bp_level, y=mvglm_sig/perma_sig, col=factor(effsize), group=effsize)) +
  geom_line() +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="relative power (mvglm/permanova)",
       title = "relative power of tests on total biomass",
       subtitle = "25% coverage; 10, 25, or 40% reduction on catch of species 2 when observed",
       col="observer effect size"
  ) +
  scale_x_continuous(labels=scales::label_percent())
```




```{r, fig.width=8}
bootstrap_size = 100
n_bootstraps = 500

res_oe %>%
  mutate(t_test_sig = ifelse(tp < 0.05, 1, 0),
           f_test_sig = ifelse(Fp < 0.05, 1, 0),
           dAIC_sig = ifelse(glmmconv==1, ifelse(AICd > 2, 1, 0), NA),
           KS_test_sig = ifelse(KSp < 0.05, 1, 0),
           median_test_sig = ifelse(ci_lo > 0 | ci_hi < 0, 1, 0),
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  select(effsize, bp_level, ends_with("sig")) %>%
  mutate(bin = as.factor(as.numeric(bp_level))) %>%
  split(~bin + effsize) %>%
  map(.x=., ~{
      slice_sample(.x, n=bootstrap_size*n_bootstraps, replace=TRUE) %>%
      mutate(boot = rep(1:n_bootstraps, each=bootstrap_size)) %>%
      group_by(bin, effsize, boot) %>%
      summarize(across(ends_with("sig"), mean), 
                .groups="drop")
    }) %>%
  list_rbind() %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni. GLMM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  group_by(bin, effsize, test) %>%
  summarize(meansig = mean(pctsig), lo = quantile(pctsig,0.025), hi = quantile(pctsig, 0.975), .groups="drop") %>%
  filter(test %in% c('MV GLM','permanova','uni. GLMM')) %>%
  mutate(bp_level = as.numeric(as.character(bin))) %>%
  #ggplot(aes(x=bp_level, y=meansig, ymin=lo, ymax=hi, col=test, fill=test, group=test)) +
  ggplot(aes(x=effsize/100, y=meansig, ymin=lo, ymax=hi, col=test, fill=test, group=test)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_ribbon(alpha = 0.2) +
  scale_color_brewer(palette = 'Set1') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  labs(#x="% dominance by species 1",
      x = "observer effect size",
       y="% of simulations with positive (significant) result",
       title = "power of tests on total biomass",
       subtitle = "25% coverage; 10, 25, or 40% reduction on catch of species 2 when observed"
  ) +
  scale_x_continuous(labels=scales::label_percent()) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1)) +
  #facet_wrap(~effsize)
  facet_wrap(~bp_level)
```










































```{r}
set.seed(123)
# Set target coverage rate
trip_coverage <- 0.25
# Set bias levels for species (change on observed trips; 0 = no bias, -0.25 = 25% reduction)
bias <- c(0, -0.05)

trip_sets_adj2 <- map(trip_sets, ~{.x %>%
  mutate(obs = rbinom(nrow(.), 1, trip_coverage)) %>%
  mutate(sp_1 = ifelse(obs == 1, sp_1 * (1+bias[1]), sp_1),
         sp_2 = ifelse(obs == 1, sp_2 * (1+bias[2]), sp_2),
         biomass_total = sp_1 + sp_2
         )
}, .progress = TRUE)

bias <- c(0, -0.5)

trip_sets_adj3 <- map(trip_sets, ~{.x %>%
  mutate(obs = rbinom(nrow(.), 1, trip_coverage)) %>%
  mutate(sp_1 = ifelse(obs == 1, sp_1 * (1+bias[1]), sp_1),
         sp_2 = ifelse(obs == 1, sp_2 * (1+bias[2]), sp_2),
         biomass_total = sp_1 + sp_2
         )
}, .progress = TRUE)
```




















































simulated data does not have block or additional variable 
```{r calculate_stats}
system.time(r <- ObserverEffectStats(trip_sets_adj[[1]], metrics = c("sp_1", "sp_2", "biomass_total"),
                      nperm_mvglm = 100, nboot_triplet = 30, nperm_adon = 30, mv_block = "vessnum")); r
#10, 30, 30 = 67s
#10, 30, 300 = 80s
#10, 300, 30 = 71s
#10, 300, 300 = 83s
#100, 30, 30 = 608s
#same but 500 trips instead of 1000 = 263s 
#before chatgpt replace = 336s (500, 100, 30, 30); after = 185s (mostly from qtweedie improvement)
  #6s for np = 3, 53s for np = 30, 247s for np = 100


set.seed(123)
stats_sets <- map(trip_sets_adj, ~{
  ObserverEffectStats(.x, metrics = c("sp_1", "sp_2", "biomass_total"),
                      nperm_mvglm = 200, nboot_triplet = 300, nperm_adon = 300,
                      #mv_block = "vessnum"
                      )
}, .progress = TRUE)
#6:47, 
now()

save(stats_sets, file="output_data/stat_sets_20251211.Rdata")
```


```{r}
stats_sets %>%
  list_rbind(names_to = "set") %>%
  mutate(bp_true = sum_O/(sum_O + sum_U)) %>%
  select(set, metric, mvglm_p, perma_p, tp)

map(trip_sets_adj, ~{
        .x %>%
          summarize(bp_level = max(bp_level), bp_true = max(bp_true), cov = mean(obs))
    }) %>%
  list_rbind(names_to = "set") %>%
  left_join (
    stats_sets %>%
      list_rbind(names_to = "set")
    , by = "set"
  ) %>%
  assign_sig() %>%
  filter(!is.na(metric)) %>%
  group_by(bp_level, metric) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  ggplot(aes(x=bp_level, y=pctsig, col=test)) +
  geom_line() +
  facet_wrap(~metric) +
  labs(title = "30 reps, 50% bias on sp2")
```







```{r}
system.time(r <- trip_sets_adj[[3]] %>%
      mutate(observed = ifelse(obs==1, 'Y', 'N')) %>%
      pivot_longer(cols=starts_with("sp_"),
                   names_to = 'species', values_to = 'biomass') %>%
      mvglm_obs(block = NULL, add_var = NULL, n_permutations=1,
                xi.vec = seq(1.2, 1.8, length.out=5))
)

system.time(r <- trip_sets_adj[[3]] %>%
      mutate(observed = ifelse(obs==1, 'Y', 'N')) %>%
      pivot_longer(cols=starts_with("sp_"),
                   names_to = 'species', values_to = 'biomass') %>%
      mvglm_obs_dd(block = NULL, add_var = NULL, n_permutations=1,
                xi.vec = seq(1.2, 1.8, length.out=5))
)
```






