---
title: "R Notebook"
output: html_notebook
---

Problem statement: species do not always have constant observer effects - relative biomass can mask effects on non-dominant species.

Goal: test a new method to do multivariate analysis on species catches and compare these results to those derived from alternative multivariate and single species or univariate approaches.  We do this over multiple degrees of species interactions (dominance) and degrees of bias (observer effects). 
this is a test

#setup

```{r packages}
library(tidyverse)
library(gdrive)       # devtools::install_github("noaa-afsc/gdrive")
library(mvobservr)    # devtools::install_github("noaa-afsc/mvobservr")
```

##combine datasets
```{r}

# Set the destination folder for data outputs on the Google shared drive
output_dribble <- gdrive_set_dribble(folder_id = "1Wh-ZQlJ3AIVaQZTWk4QNuyiMfoVECQgt")

#bias 25
load(gdrive_download("output_data/alltests_2spp_oe25_cov25_set1.Rdata", output_dribble))
r1 <- res_comb
load(gdrive_download("output_data/alltests_2spp_oe25_cov25_set2.Rdata", output_dribble))
r2 <- res_comb %>% mutate(set = set + max(r1$set))
res_comb <- bind_rows(r1, r2)
rm(r1, r2); gc() #remove temp objects and clean up
res_oe <- res_comb %>% mutate(effsize = 25)

#bias 10
load(gdrive_download("output_data/alltests_2spp_oe10_cov25_set1.Rdata", output_dribble))
r1 <- res_comb
load(gdrive_download("output_data/alltests_2spp_oe10_cov25_set2.Rdata", output_dribble))
r2 <- res_comb %>% mutate(set = set + max(r1$set))
res_comb <- bind_rows(r1, r2)
rm(r1, r2); gc() #remove temp objects and clean up
res_oe <- res_oe %>%
  bind_rows(res_comb %>% mutate(effsize = 10))

#bias 40
load(gdrive_download("output_data/alltests_2spp_oe40_cov25_set1.Rdata", output_dribble))
r1 <- res_comb
load(gdrive_download("output_data/alltests_2spp_oe40_cov25_set2.Rdata", output_dribble))
r2 <- res_comb %>% mutate(set = set + max(r1$set))
res_comb <- bind_rows(r1, r2)
rm(r1, r2); gc() #remove temp objects and clean up
res_oe <- res_oe %>%
  bind_rows(res_comb %>% mutate(effsize = 40))

table(res_oe$effsize, res_oe$bp_level)
```


##figure: relative power for mvglm and permanova vs dominance for 3 levels of bias
```{r}
res_oe %>%
  mutate(
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(bp_level, effsize) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  ggplot(aes(x=bp_level, y=mvglm_sig/perma_sig, col=factor(effsize), group=effsize)) +
  geom_line() +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="relative power (mvglm/permanova)",
       title = "relative power of tests on total biomass",
       subtitle = "25% coverage; 10, 25, or 40% reduction on catch of species 2 when observed",
       col="observer effect size"
  ) +
  scale_x_continuous(labels=scales::label_percent())
```



##figure: power of 3 tests vs bias w/ bootstrap error ribbons faceted by dominance
```{r, fig.width=8}
bootstrap_size = 100
n_bootstraps = 500

res_oe %>%
  mutate(t_test_sig = ifelse(tp < 0.05, 1, 0),
           f_test_sig = ifelse(Fp < 0.05, 1, 0),
           dAIC_sig = ifelse(glmconv==1, ifelse(AICd > 2, 1, 0), NA),
           KS_test_sig = ifelse(KSp < 0.05, 1, 0),
           median_test_sig = ifelse(ci_lo > 0 | ci_hi < 0, 1, 0),
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  select(effsize, bp_level, ends_with("sig")) %>%
  mutate(bin = as.factor(as.numeric(bp_level))) %>%
  split(~bin + effsize) %>%
  map(.x=., ~{
      slice_sample(.x, n=bootstrap_size*n_bootstraps, replace=TRUE) %>%
      mutate(boot = rep(1:n_bootstraps, each=bootstrap_size)) %>%
      group_by(bin, effsize, boot) %>%
      summarize(across(ends_with("sig"), mean), 
                .groups="drop")
    }) %>%
  list_rbind() %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni. GLM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  group_by(bin, effsize, test) %>%
  summarize(meansig = mean(pctsig), lo = quantile(pctsig,0.025), hi = quantile(pctsig, 0.975), .groups="drop") %>%
  filter(test %in% c('MV GLM','permanova','uni. GLM')) %>%
  #mutate(bp_level = paste0("BP: ", as.numeric(as.character(bin))*100, "%")) %>%
  mutate(bp_level = as.numeric(as.character(bin))) %>%
  ggplot(aes(x=bp_level, y=meansig, ymin=lo, ymax=hi, col=test, fill=test, group=test)) +
  #ggplot(aes(x=effsize/100, y=meansig, ymin=lo, ymax=hi, col=test, fill=test, group=test)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_ribbon(alpha = 0.2) +
  scale_color_brewer(palette = 'Set1') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  labs(#x="% dominance by species 1",
      x = "observer effect size",
       y="% of simulations with positive (significant) result",
       title = "power of tests on total biomass",
       subtitle = "25% coverage; 10, 25, or 40% reduction on catch of species 2 when observed \n 200 reps per data point"
  ) +
  scale_x_continuous(labels=scales::label_percent()) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1)) +
  facet_wrap(~effsize)
  #facet_wrap(~bp_level)
```

##figure: power of all tests for 3 bias levels
```{r}
res_oe %>%
  mutate(t_test_sig = ifelse(tp < 0.05, 1, 0),
           f_test_sig = ifelse(Fp < 0.05, 1, 0),
           permu_sig = ifelse(p_val < 0.05, 1, 0),
           dAIC_sig = ifelse(glmconv==1, ifelse(AICd > 2, 1, 0), NA),
           KS_test_sig = ifelse(KSp < 0.05, 1, 0),
           median_test_sig = ifelse(ci_lo > 0 | ci_hi < 0, 1, 0),
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(effsize, bp_level) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "permu_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni.permu", "uni. GLM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  mutate(effsize = paste0(effsize, "% observer effect")) %>%
  ggplot(aes(x=bp_level, y=pctsig, col=test)) +
  geom_line(linewidth=1) +
  scale_color_brewer(palette = 'Set2') +
  theme_bw() +
  labs(x="% dominance by species 1",
       y="% of simulations with positive (significant) result",
       title = "power of tests on total biomass",
       subtitle = "25% coverage, reduction on catch of species 2 when observed, 200 reps"
  ) +
  scale_x_continuous(labels=scales::label_percent()) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1)) +
  facet_wrap(~effsize)
```