---
title: "investigate scalar"
output: html_notebook
---

```{r}
library(tidyverse)
library(FMAtools)  
library(mvobservr)
library(glmmTMB)
library(tweedie)
```


```{r fixedvar}
tweedie_power <- 1.6
phi <- 3
fixed_total_biomass <- 1000000
ntrips <- 500
target_bp_level = 0.6
```


```{r generatedata}
mu_scalar = 100

catches <- data.frame(uid = 1:ntrips)
catches$sp_1 <- rtweedie(ntrips, p = tweedie_power, mu = mu_scalar*target_bp_level, phi = phi)
catches$sp_2 <- rtweedie(ntrips, p = tweedie_power, mu = mu_scalar*(1-target_bp_level), phi = phi)

#standardize catches
catches <- catches %>%
  mutate(total_biomass = sum(sp_1 + sp_2)) %>%
  mutate(scalar = fixed_total_biomass/total_biomass) %>%
  mutate(across(c(sp_1, sp_2), function(x) x*scalar)) %>%
  select(-scalar, -total_biomass) %>%
  mutate(biomass_total = sp_1 + sp_2)

```

check that biomass total and bp are correct
```{r}
catches %>% 
  summarize(sum(biomass_total), sum(sp_1)/sum(biomass_total))
```
look at mean, sd, and proprotion of 0s of sp1 and sp2
```{r}
catches %>%
  summarize(mean(sp_1), sd(sp_1), mean(sp_1==0), mean(sp_2), sd(sp_2), mean(sp_2==0))
```

look at distributions
```{r}
catches_piv <- catches %>%
  pivot_longer(cols=c(sp_1,sp_2,biomass_total)) 

catches_piv_agg <- catches_piv %>%
  group_by(name) %>%
  summarize(mean = mean(value), sd = sd(value))

catches_piv %>%
  ggplot(aes(x=value)) +
  geom_histogram(bins=50, col="grey30", fill="grey80") +
  facet_wrap(~name) +
  theme_bw() +
  geom_vline(data=catches_piv_agg, aes(xintercept = mean), col='blue') +
  geom_vline(data=catches_piv_agg, aes(xintercept = mean-sd), col='blue', linetype='dashed') +
  geom_vline(data=catches_piv_agg, aes(xintercept = mean+sd), col='blue', linetype='dashed')
```

confirm tweedie params
```{r}
tw_fit_1 <- glmmTMB(sp_1 ~ 1, family=tweedie(), data=catches)
tw_fit_1

tw_fit_2 <- glmmTMB(sp_2 ~ 1, family=tweedie(), data=catches)
tw_fit_2
```

power estimate = xi; set tweedie_power at 1.6
dispersion parameter = phi; set phi at 3
intercept = log(mu); set mu at 60 (sp_1) and 40 (sp_2)



without scaling to a fixed total biomass, would expect to see
p_est ~ 1.6
phi_est ~ 3
mu_est ~ 0.6*mu_scalar (sp_1) or 0.4*mu_scalar (sp_2)

```{r}
p_est   <- family_params(tw_fit_1)  # Tweedie power (p)
phi_est <- sigma(tw_fit_1)          # Dispersion (phi)
mu_est  <- exp(fixef(tw_fit_1)$cond["(Intercept)"]) # Mean (mu)

cat("results for sp_1\nPower (p):", p_est, "\nDispersion (phi):", phi_est, "\nMean (mu):", mu_est)
```
```{r}
p_est   <- family_params(tw_fit_2)  # Tweedie power (p)
phi_est <- sigma(tw_fit_2)          # Dispersion (phi)
mu_est  <- exp(fixef(tw_fit_2)$cond["(Intercept)"]) # Mean (mu)

cat("results for sp_2\nPower (p):", p_est, "\nDispersion (phi):", phi_est, "\nMean (mu):", mu_est)
```


#comparing across mu_scalar values
```{r}
muTest <- function(mu_scalar = 1, scale_to_fixed_biomass = FALSE) {
  catches <- data.frame(uid = 1:ntrips)
  catches$sp_1 <- rtweedie(ntrips, p = tweedie_power, mu = mu_scalar*target_bp_level, phi = phi)
  catches$sp_2 <- rtweedie(ntrips, p = tweedie_power, mu = mu_scalar*(1-target_bp_level), phi = phi)
  
  if(scale_to_fixed_biomass) {
    #standardize catches
    catches <- catches %>%
      mutate(total_biomass = sum(sp_1 + sp_2)) %>%
      mutate(scalar = fixed_total_biomass/total_biomass) %>%
      mutate(across(c(sp_1, sp_2), function(x) x*scalar)) %>%
      select(-scalar, -total_biomass) 
  } 
  
  catches <- catches %>%
    mutate(biomass_total = sp_1 + sp_2)
  
  catch_summary <- catches %>% 
    summarize(total_biomass = sum(biomass_total), true_bp = sum(sp_1)/sum(biomass_total),
              mean_sp1 = mean(sp_1), sd_sp1 = sd(sp_1), zeros_sp1 = mean(sp_1==0), 
              mean_sp2 = mean(sp_2), sd_so2 = sd(sp_2), zeros_sp2 = mean(sp_2==0))
  
  tw_fit_sp1 <- glmmTMB(sp_1 ~ 1, family=tweedie(), data=catches)
  tw_fit_sp2 <- glmmTMB(sp_2 ~ 1, family=tweedie(), data=catches)
  
  p_est_sp1   <- family_params(tw_fit_sp1)  # Tweedie power (p)
  phi_est_sp1 <- sigma(tw_fit_sp1)          # Dispersion (phi)
  mu_est_sp1  <- exp(fixef(tw_fit_sp1)$cond["(Intercept)"]) # Mean (mu)
  
  p_est_sp2   <- family_params(tw_fit_sp2)  # Tweedie power (p)
  phi_est_sp2 <- sigma(tw_fit_sp2)          # Dispersion (phi)
  mu_est_sp2  <- exp(fixef(tw_fit_sp2)$cond["(Intercept)"]) # Mean (mu)
  
  fit_summary <- data.frame(p_est_sp1, phi_est_sp1, mu_est_sp1, p_est_sp2, phi_est_sp2, mu_est_sp2) %>%
    remove_rownames()
  
  return(cbind(catch_summary, fit_summary))
}
```

```{r}
muTest(mu_scalar = 100, scale_to_fixed_biomass = FALSE)
```
fixed_total_biomass/ntrips = 2000
```{r}
mu_results_notscaled <- map(rep(c(1,10,100,1000,2000,5000),50), ~{
  muTest(mu_scalar = .x) %>%
    mutate(mu_scalar = .x, .before=1)
}, .progress=TRUE) %>%
  list_rbind()

mu_results_notscaled %>%
  pivot_longer(cols=-mu_scalar) %>%
  ggplot(aes(x=mu_scalar, y=value)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~name, scales="free") +
  theme_bw()
```
without scaling to a fixed total biomass, would expect to see
p_est ~ 1.6
phi_est ~ 3
mu_est ~ 0.6*mu_scalar (sp_1) or 0.4*mu_scalar (sp_2)

```{r}
mu_results_notscaled %>%
  group_by(mu_scalar) %>%
  summarize(across(c(p_est_sp1, p_est_sp2, phi_est_sp1, phi_est_sp2, mu_est_sp1, mu_est_sp2, zeros_sp1, zeros_sp2), mean))
```

p_est (power) unchanged by mu_scalar
phi_est (dispersion) unchanged at mu_scalar 1, 10, 100 but increases after that
mu_est (mean) changes as expected with scalar - keeps 0.6 or 0.4 * mu_scalar

It's in the 0s that most of the differences occur. 
At mu_scalar = 1, about 50% of the trips will have a 0 value
At mu_scalar = 10, about 20%
At mu_scalar = 100, about 2%

this is without total_biomass scaling (making the fleet total a fixed amount) but that wouldn't affect the rate of 0s


#compare mvglm performance for mu_scalar values
```{r}
# Set how many populations per scenario to generate
n_samples_per_level <- 10
# Set the Tweedie power parameter (lambda). 1 < p < 2 is typical for biomass.
tweedie_power <- 1.6
# Set the dispersion parameter for the Tweedie distribution.
# Higher values create more variance (further from target BP).
phi <- 3

# Set the desired total biomass for every saved population.
fixed_total_biomass <- 1000000
# Set target Berger-Parker index (dominance) levels.
fixed_bp_level <- 0.6


# Set the number of trips
target_mu_scalar_levels <- c(1, 10, 100, 1000)

# total datasets to create
n_samples_per_level * length(target_mu_scalar_levels)
```

```{r generatedata}
# for reproducibility
set.seed(123)

trip_sets <- map(rep(1:length(target_mu_scalar_levels), n_samples_per_level), ~{
  #set up data frame for trips
  catches <- data.frame(uid = 1:ntrips)
  
  #create catches
  catches$sp_1 <- rtweedie(ntrips, p = tweedie_power, mu = target_mu_scalar_levels[.x]*fixed_bp_level, phi = phi)
  catches$sp_2 <- rtweedie(ntrips, p = tweedie_power, mu = target_mu_scalar_levels[.x]*(1-fixed_bp_level), phi = phi)
  # the 100 acts as a raising factor to get away from a lot of 0s
  
  #standardize catches
  catches <- catches %>%
    mutate(total_biomass = sum(sp_1 + sp_2)) %>%
    mutate(scalar = fixed_total_biomass/total_biomass) %>%
    mutate(across(c(sp_1, sp_2), function(x) x*scalar)) %>%
    select(-scalar, -total_biomass) %>%
    mutate(biomass_total = sp_1 + sp_2) %>% #this is at the trip level, previously was at the fleet level
    mutate(bp_level = fixed_bp_level,
           bp_true = sum(sp_1)/sum(biomass_total),
           mu_scalar = target_mu_scalar_levels[.x])
  
}, .progress = TRUE)

```



```{r addbias}
# Set target coverage rate
trip_coverage <- 0.25
# Set bias levels for species (change on observed trips; 0 = no bias, -0.25 = 25% reduction)
bias <- c(0, -0.25)
#bias <- c(0,0) #no bias (for false positives)

trip_sets_adj <- map(trip_sets, ~{.x %>%
  mutate(obs = rbinom(nrow(.), 1, trip_coverage)) %>%
  mutate(sp_1 = ifelse(obs == 1, sp_1 * (1+bias[1]), sp_1),
         sp_2 = ifelse(obs == 1, sp_2 * (1+bias[2]), sp_2),
         biomass_total = sp_1 + sp_2
         )
}, .progress = TRUE)
```


##permanova
```{r}
res_p <- map(trip_sets_adj, ~{
  Y <- .x %>% dplyr::select(starts_with("sp_"))
  adon <- vegan::adonis2(Y~factor(obs), data = .x, permutations = 1000, method="euclidean")
  data.frame(perma_p = adon$`Pr(>F)`[1])
}, .progress=TRUE) #20min for 500 (was about 35 on laptop)
res_p <- list_rbind(res_p, names_to = "set")
res_p
```

##mvglm
```{r}
res_m <- imap(1:length(trip_sets_adj), ~{
  print(paste0("***set ", .x, " ", Sys.time()))
  r <- trip_sets_adj[[.x]] %>%
      mutate(observed = ifelse(obs==1, 'Y', 'N')) %>%
      pivot_longer(cols=starts_with("sp_"),
                   names_to = 'species', values_to = 'biomass') %>%
      mvglm_obs(block = NULL, add_var = NULL, n_permutations=100,
                xi.vec = seq(1.2, 1.8, length.out=5),
                nCores=12)
  data.frame(mvglm_p = r$results$p, sp_1 = r$results$uni.p[1], sp_2 = r$results$uni.p[2])
      
}) #about an hour for 40 trip_sets
res_m <- list_rbind(res_m, names_to = "set") 
res_m

save(res_m, file="output_data/mvglmloop_mu_scalar.Rdata")
```



```{r}
res_comb <- map(trip_sets_adj, ~{
        .x %>%
          summarize(mu_scalar = mean(mu_scalar), cov = mean(obs), sp1 = sum(sp_1), sp2 = sum(sp_2))
    }) %>%
  list_rbind(names_to = "set") %>%
  left_join(res_p, by="set") %>%
  left_join(res_m, by="set")
res_comb %>%
  mutate(b = fixed_total_biomass-(sp1+sp2),
         bpct = b/sp2/cov) #check to make sure bias is being applied correctly

save(res_comb, file="rescomb_mu_scalar.Rdata")
```


```{r}
res_comb %>%
  #assign_sig() %>%
  mutate(
           mvglm_sig = ifelse(mvglm_p < 0.05, 1, 0),
           perma_sig = ifelse(perma_p < 0.05, 1, 0)
    ) %>%
  group_by(mu_scalar) %>%
  summarize(across(ends_with("sig"), ~mean(.x, na.rm=TRUE)), .groups="drop") %>%
  pivot_longer(cols = ends_with("sig"), names_to = "test", values_to = "pctsig") %>%
  mutate(test = factor(test, levels = c("t_test_sig", "f_test_sig", "dAIC_sig","KS_test_sig","median_test_sig","mvglm_sig", "perma_sig"),
                       ordered = TRUE,
                       labels = c("t-test", "F test", "uni. GLMM", "K-S test", "med. diff.", "MV GLM", "permanova"))) %>%
  ggplot(aes(x=as.character(mu_scalar), y=pctsig, col=test, group=test)) +
  geom_line() +
  theme_bw() +
  labs(x="mu_scalar",
       y="% of simulations with positive (significant) result",
       title = "power vs # of trips",
       subtitle = "60% dominance, 10 trip sets per level"
  ) +
  scale_y_continuous(labels=scales::label_percent(), limits=c(0,1)) 
```


```{r}
res_comb %>%
  select(mu_scalar, set, mvglm_p, sp_1, sp_2) %>% #sp_1 and sp_2 here are the uni.p outputs from mvglm
  arrange(mu_scalar, mvglm_p)
```
at mu_scalar 1 or 10, even the uni.p output from mvglm for sp_2 is not always detecting a difference

look at one tripset
```{r}
trip_sets_adj[[26]] %>%
  ggplot(aes(x=sp_2, fill=factor(obs))) +
  geom_histogram(position='dodge')

trip_sets_adj[[26]] %>%
  group_by(obs) %>%
  summarize(q10 = quantile(sp_2,0.1),
            q25 = quantile(sp_2,0.25),
            q50 = quantile(sp_2,0.5),
            q75 = quantile(sp_2,0.75),
            q90 = quantile(sp_2,0.9)
            )
```








